---
title: "Enron Report"
author: "By: Courtney Ferguson Lee"
date: "June 29, 2017"
output: html_document
---

<style type="text/css">
body{ /* Normal  */
      font-size: 18px;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F,
                      fig.align = 'center')
```

```{r Load_Packages}
library(tidyverse)
library(ggplot2)
library(scales) # need to get labels = comma
library(GGally)
```

```{r Load_Data}
setwd('/Users/courtneyfergusonlee/p5')
enron <- read.csv('enron_data.csv', 
                  na.strings = 'NaN',
                  stringsAsFactors = F)
```

```{r Reshape_Data}
enron.new_features <- enron %>%
  mutate(fract_to_poi = from_this_person_to_poi/from_messages,
         fract_from_poi = from_poi_to_this_person/to_messages,
         exer_stock_opts_over_tot = exercised_stock_options/total_stock_value,
         poi = as.factor(poi))
```



> 1. Summarize for us the goal of this project and how machine learning is 
useful in trying to accomplish it. As part of your answer, give some background 
on the dataset and how it can be used to answer the project question. Were there
any outliers in the data when you got it, and how did you handle those?  
[relevant rubric items: “data exploration”, “outlier investigation”]

The goal for this project was to identify persons of interest (POIs) from a 
dataset of enron employees.  The dataset initially contained 146 observations 
across 21 features and a POI status label.  There were 18 POI's and 128 
non-POI's.  I used a supervised machine learning algorithm to identify patterns 
in the email and financial data that separated POIs from non-POIs. The financial
and email variables included:

Financial Variables | Email Variables
---|---
bonus | email_address
deferral_payments | from_messages
deferred_income | from_poi_to_this_person
director_fees | from_this_person_to_poi
exer_stock_opts_over_tot | shared_receipt_with_poi
exercised_stock_options | to_messages
expenses | 
loan_advances | 
long_term_incentive | 
other | 
restricted_stock | 
restricted_stock_deferred | 
salary | 
salary_over_bonus | 
total_payments | 
total_stock_value | 

```{r}
subset(enron, is.na(total_payments)
       & is.na(total_stock_value)
       & is.na(email_address))
```

  

Three outliers in the dataset needed to be removed: “TOTAL”, “THE TRAVEL AGENCY 
IN THE PARK”, and "LOCKHART EUGENE E".  The first two weren't actual people, 
so they did not contribute to the model.  Eugene Lockhart was missing data for
all of his features, so he was removed as well.  There were several financial 
and email outliers, but I ultimately included them because they contained useful
information.  Because of the nature of the dataset, it was difficult to remove 
any outliers that varied significantly from the rest of the population.  For 
example, Ken Lay and Joseph Hirko had the largest exercised stock options, but 
they were both persons of interest.  In addition, the dataset only consisted of 
less than 150 people and there was a lot of missing data, as shown in the 
following chart.  With such a small number of observations, excluding any 
outliers would have had a significant impact on the model.


```{r Missing_Values_Heatmap}
enron.missing <- as.data.frame(sapply(enron.new_features, is.na))

enron.long <- enron.new_features %>%
  gather(var_name,
         value,
         -name,
         na.rm = F)

enron.long.missing <- enron.long %>%
  mutate(value = !is.na(value))

ggplot(data = enron.long.missing,
       aes(x = name,
           y = var_name)) +
  geom_tile(aes(fill = value)) +
  scale_fill_manual(values = c('tomato2', 'springgreen3'),
                    labels = c('Missing', 'Present')) +
  #scale_fill_brewer(palette = "RdYlGn") +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  ggtitle("Heatmap of Missing Values") +
  xlab("Observations") +
  ylab("Variable Name") +
  labs(fill = "Data Availability")
```



> 2. What features did you end up using in your POI identifier, and what 
selection process did you use to pick them? Did you have to do any scaling? Why 
or why not? As part of the assignment, you should attempt to engineer your own 
feature that does not come ready-made in the dataset -- explain what feature you
tried to make, and the rationale behind it. (You do not necessarily have to use
it in the final analysis, only engineer and test it.) In your feature selection
step, if you used an algorithm like a decision tree, please also give the feature 
importances of the features that you use, and if you used an automated feature 
selection function like SelectKBest, please report the feature scores and 
reasons for your choice of parameter values.  [relevant rubric items: “create 
new features”, “intelligently select features”, “properly scale features”]

Variable | Definition
-------------- | ------------------------------------------------
fract_from_poi | Fraction of total emails received from a poi
fract_to_poi | Fraction of total emails sent to a poi
director_fees | Compensation for attendance at board meetings
restricted_stock_deferred | Stock that was not fully transferrable until certain contitions were met
exercised_stock_options | Amount of Enron stock an employee bought or sold
expenses |  Costs occurred by an employee while conducting business

I used Exploratory Data Analysis to identify / engineer the above 6 features to 
help create clear boundaries for my classifier. I created two of the features: 
the fraction of emails sent to POIs and the fraction of emails from POIs.  I 
did this because it did not make sense to use the raw number of emails to or 
from a POI.  The total number of sent and received emails ranged from around 10 
to over 10,000.  To better compare human interactions, it made more sense to 
code this data as a proportion.  When I created the scatterplot of the two 
features (first plot below), they created clear boundaries that helped improve 
my decision tree classifier.  The fraction of emails sent to a POI (x-axis)
below .16 and above .68 uniformly contained non-POI's. The same was true for the
fraction of emails from a POI (y-axis) below .02 and above .14. The original 
variables yielded the following evaluation metrics:

<img src = 'screenshots/Screen Shot 2017-07-03 at 12.21.23 PM.png'>

When I included just the above 6 variables, the numbers increased:

<img src = 'screenshots/Screen Shot 2017-07-03 at 12.21.56 PM.png'>

The plots below helped me understand the underlying nature of the financial and
email variables:


```{r}
ggplot(data = enron.new_features,
       aes(x = fract_to_poi,
           y = fract_from_poi,
           color = poi)) +
  geom_point() +
  scale_x_continuous(breaks = seq(0, 1.00, .04)) +
  scale_y_continuous(breaks = seq(0, .23, .01)) +
  ggtitle('Fraction of Emails From POI vs Fraction of Emails to POI') +
  xlab('Fraction Sent to POI') +
  ylab('Fraction From POI') +
  labs(color = 'POI')
```


```{r Financial_long_form}
financial_long <- enron %>%
  mutate(deferred_income = -deferred_income,
         restricted_stock_deferred = -restricted_stock_deferred) %>%
  select(expenses,
         other,
         director_fees,
         restricted_stock_deferred,
         deferred_income,
         deferral_payments,
         salary,
         long_term_incentive,
         restricted_stock,
         bonus,
         total_payments,
         total_stock_value,
         exercised_stock_options,
         loan_advances,
         poi) %>%
  gather(financial_var,
         amount,
         -poi) %>%
  mutate(financial_var = as.factor(financial_var))
```


```{r Financial_Boxplot}
ggplot(data = subset(financial_long, 
                     !is.na(amount) & amount > 0),
       aes(x = reorder(financial_var, 
                       amount, 
                       median),
           y = amount)) +
  geom_boxplot() +
  geom_jitter(aes(color = poi,
                  alpha = .2),
              width = .2,
              size = .7) +
  scale_y_log10(labels = comma,
                breaks = 10^seq(2, 8, 1)) +
  theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
  ggtitle("Distribution of Financial Variables") +
  xlab("Financial Variable") +
  ylab("Amount") +
  guides(alpha = 'none') +
  labs(color = 'POI')

```

```{r Email_Long_Form}
email_long <- enron %>%
  select(to_messages,
         shared_receipt_with_poi,
         from_messages,
         from_this_person_to_poi,
         from_poi_to_this_person,
         poi) %>%
  gather(email_feature,
         amount,
         -poi) %>%
  mutate(email_feature = as.factor(email_feature))
```

```{r Email_Boxplot}
ggplot(data = subset(email_long, !is.na(amount)),
       aes(x = reorder(email_feature, amount, median),
           y = amount + 1)) +
  geom_boxplot() +
  geom_jitter(aes(alpha = .2,
                  color = poi),
              width = .2) +
  scale_y_log10(labels = comma,
                breaks = c(0,
                           10,
                           100,
                           1000,
                           10000)) +
  ggtitle('Distribution of Emails') +
  xlab('Email Feature') +
  ylab('Number of Messages') +
  guides(alpha = 'none') +
  labs(color = 'POI')
```


```{r Email_Scatter}
ggplot(data = enron.new_features,
       aes(x = fract_to_poi,
           y = fract_from_poi,
           color = poi)) +
  geom_point() +
  ggtitle('Fraction of Emails From POI vs Fraction of Emails to POI') +
  xlab('Fraction Sent to POI') +
  ylab('Fraction From POI') +
  labs(color = 'POI')
```

After identifying useful features, I scaled them before using PCA in my model. I
scaled the features because the ranges of the variables were too large and would
throw off predictions.  For instance, the range for fractional email variables 
ranged from 0.0 to 1.0 while exercised stock options varied from 1,000 to 
around 30,000,000.  With such a large range, PCA would not be able to properly
determine which variable accounted for most of the variance in the data. I 
created my own MinMaxScaler and applied it right after creating new features.
It ensured that all variables ranged from 0 to 1.  PCA was a necessary feature
selection step because it allowed the data to speak for itself in a way. I also
used GridSearchCV to tune the parameters of PCA itself.  The algorithm tested
a range of 2 to 6 components to determine the ideal number to use. It identified
that 6 was the ideal number. The explained variance ratios from PCA and 
features importances from DecisionTree were: 

**Explained Variance Ratios:**
[ 0.3574516,   0.23236625,  0.20634216,  0.11222562,  0.08776709]

  
**Feature Importances:**
[ 0.,          0.,          0.15539112,  0.57452213,  0.27008675]


> 3. What algorithm did you end up using? What other one(s) did you try? How did 
model performance differ between algorithms?  [relevant rubric item: “pick an algorithm”]

I tested out several algorithms before settling on a decision tree. I started 
with a GaussianNaiveBayes to get a benchmark, then tried out Support Vector
Machines, KNearest Neighbors, AdaBoost and finally DecisionTree. The KNearest 
Neighbors Classifier had the worst results overall at only .22 for precision, 
recall and F1 scores.  The Support Vector Machines model but was slower than the
others and didn't improve the F1 scores much.  AdaBoost had similar results. At
that point, I decided to try the DecisionTree Classifier and added PCA to 
transform the variables.  This performed the best overall in terms of both 
evaluation metrics and speed.  The original survey results without applying PCA
are summarized below.  I used the following features in each model:

Classifier | Results
----- | -----
DecisionTree | <img src = 'screenshots/Screen Shot 2017-07-03 at 12.21.56 PM.png'>
Naive Bayes | <img src = 'screenshots/Screen Shot 2017-06-21 at 5.48.50 PM.png'>



> 4. What does it mean to tune the parameters of an algorithm, and what can 
happenif you don’t do this well?  How did you tune the parameters of your 
particularalgorithm? What parameters did you tune? (Some algorithms do not have
parameters that you need to tune -- if this is the case for the one you picked,
identify and briefly explain how you would have done it for the model that was 
not your final choice or a different model that does utilize parameter tuning, 
e.g. a decision tree classifier).  [relevant rubric items: “discuss parameter 
tuning”, “tune the algorithm”]

Tuning the parameters of an algorithm involves reading the sklearn documentation
and trying out different combinations of input arguments.  For instance, when I
settled on a DecisionTree Classifier, I played around with the max_depth,
minimal_samples_split, and criterion.  We calculated entropy for various
scenarios in the lessons so I tried using that as an alternative criterion, but
the results didn't change much from baseline.  I found the greatest variation in
F1 scores when I changed the PCA parameters.  When I did not specify a 
random_state, the scores kept bouncing all over the place.  I discovered that
this was an important step to ensure that others could reproduce my results.

> 5. What is validation, and what’s a classic mistake you can make if you do it 
wrong? How did you validate your analysis?  [relevant rubric items: “discuss
validation”, “validation strategy”]

Validation involves ensuring your model is effective on both training and
testing data.  One pitfall in validation is when you accidentally overfit your
model to the training data.  Overfitting is when your model performs well on
training data but poorly on new testing data.  This can happen if you don't 
partition a large enough chunk for your testing set or when you use an
overly-convoluted boundary when training.

> 6. Give at least 2 evaluation metrics and your average performance for each of 
them.  Explain an interpretation of your metrics that says something human-
understandable about your algorithm’s performance. [relevant rubric item: 
“usage of evaluation metrics”]

My average precision score was around .34 and my average recall was around .38-
.39.  Precision compares the number of false negatives to true positives in the
results.  In other words, it asks, "Out of all of the items that are truly
positive, how many were correctly labeled as positive?"  Recall measures the 
number of false positives to true positives.  It asks, "Out of all of the items
labeled as positive, how many truly belong to the positive class?"  My results
indicate that my algorith was better at lowering the number of false positives
than it was at weeding out false negatives.